---
title: "nistr - Quick Reference"
author: "Jesus Castagnetto (<jesus@castagnetto.com>)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nistr - Quick Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package implements R functions and an S3 class to access the
NIST the NIST Randomness Beacon, through its REST web service interface.

The NIST service <q>generates full-entropy bit-strings and posts them in blocks
of 512 bits every 60 seconds</q>^[https://beacon.nist.gov/home]

## Implementation details

### Functions

According to the beacon's documentation, there are five endpoints defined and
this package wraps the call to each one in a function:

```{r echo=FALSE, results='asis'}

df <- data.frame(
  Function = c(
    "`nrb_get_last()`",
    "`nrb_get_next(timestamp)`",
    "`nrb_get_previous(timestamp)`",
    "`nrb_get_record(timestamp)`",
    "`nrb_get_start_chain(timestamp)`"
  ),
  Endpoint = c(
    "`https://beacon.nist.gov/rest/record/last`",
    "`https://beacon.nist.gov/rest/record/next/<timestamp>`",
    "`https://beacon.nist.gov/rest/record/previous/<timestamp>`",
    "`https://beacon.nist.gov/rest/record/<timestamp>`",
    "`https://beacon.nist.gov/rest/record/start-chain/<timestamp>`"
  ),
  Description = c(
    "Last Record",
    "Next Record",
    "Previous Record",
    "Current Record (or next closest)",
    "Start Chain Record"
  )
      )

knitr::kable(df)
```

Where `timestamp` is the POSIX time format: the number of seconds since midnight
UTC, January 1, 1970. Alternatively a `POSIXct` o `POSIXlt` value can be used.

### Class

The service returns an XML document with several pieces of information:

- The (seed) value represented as a 64 byte (512-bit) hex string value, and its timestamp
- The SHA-512 hash value for the previous record as a 64 byte hex string
- A digital signature (RSA) computed over (in order): version, frequency, timeStamp, seedValue, previousHashValue, errorCode. *Note*: Except for version, the hash is on the byte representations and not the string representations of the data values
- The SHA-512 hash of the signature as a 64 byte hex string
- The service version and its update frequency (currenty, 60 seconds)
- The status of the response
    - 0 = Chain intact, values all good
    - 1 = Start of a new chain of values, previous hash value will be all zeroes
    - 2 = Time between values is greater than the frequency, but the chain is still intact

This package wraps this XML document in the `nrb_response` class, which contains all
the values from the beacon's response.

For example, to get the record generated on Jan 1, 2015 at 10:00h (UTC), we will
use the following code:

```{r eval=FALSE}
library(nistrb)
ts <- as.POSIXct("2015-01-01 10:00:00", tz = "UTC")
lval <- nrb_get_record(ts)
lval
```

The last line above prints:

```
NIST Randomness Beacon Response
===============================

* version:      Version 1.0
* frequency:    60 seconds

* timestamp:    2015-01-01 05:00:00 (1420106400)
* seed:         037035A8D1F2C74D6E233F6B962E7A7233BBAF67A1B0247A81
                29E0BB50D619144A4083A5375A1BC16731FFB83357DD5CC1A9
                62523370453E6707E74762B515D6
                
* prev.:        368D684FB8DC709852C0E67ECA439774AF8C7D21CFAB73083A
                1C0120C340B54C3237E1FB73C78B739884E05D8D11BA17AB9C
                F10EEA3C37CBDEB888888486BD4F
                
* signature:    20F647B95F5A56DDBEA82044821DC022EB4CC0B53700CD6351
                4CC02D320B42DAA684A1FFA9B8635E8BEE58A35DDDC899FB84
                16BD08AE57EFBF26A2325E2334F1CC5B5B0126AF0866443215
                7DE6C9FD58F857340D46DBAF902FD5D7770BFBB8755279EB92
                D4787655B5A5658DFC97EB9209D4AC271EF30D5BDBCAAE2F37
                5144BEEF5CB17A9CF37C52648ED0BC682ACFF147D2CA8CCC39
                9253B351660261A6CF80A166089EA54E7F425B652C6417735C
                2BE01414515481B05CC359836F0261CF1D7F1FC23BC0FA1A52
                125455755CDF2DEECE5D4555883897E8DCD9FB564A868A0007
                5489EA28D01949A56FD4BC5AFC14AB6766C80BECB69F31CB0D
                DFA47FA59703
                
* output:       27DADEBF52C5834E1894B99625845689C7BFC1ABD05CC672B1
                AC8A6A585424BC4C6E577C0249FBAFBDC589CD6937BEBBA764
                7E1233692D2003237598F12FF342
                
* status:       0 (Chain intact, values all good
```

