---
title: "nistr - Quick Reference"
author: "Jesus Castagnetto (<jesus@castagnetto.com>)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nistr - Quick Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package implements R functions and an S3 class to access the
NIST the NIST Randomness Beacon, through its REST web service interface.

The NIST service <q>generates full-entropy bit-strings and posts them in blocks
of 512 bits every 60 seconds</q>^[https://beacon.nist.gov/home]

## Implementation details

### Functions

According to the beacon's documentation, there are five endpoints defined and
this package wraps the call to each one in a function:

```{r echo=FALSE, results='asis'}

df <- data.frame(
  Function = c(
    "`nrb_get_last()`",
    "`nrb_get_next(timestamp)`",
    "`nrb_get_previous(timestamp)`",
    "`nrb_get_record(timestamp)`",
    "`nrb_get_start_chain(timestamp)`"
  ),
  Endpoint = c(
    "`https://beacon.nist.gov/rest/record/last`",
    "`https://beacon.nist.gov/rest/record/next/<timestamp>`",
    "`https://beacon.nist.gov/rest/record/previous/<timestamp>`",
    "`https://beacon.nist.gov/rest/record/<timestamp>`",
    "`https://beacon.nist.gov/rest/record/start-chain/<timestamp>`"
  ),
  Description = c(
    "Last Record",
    "Next Record",
    "Previous Record",
    "Current Record (or next closest)",
    "Start Chain Record"
  )
      )

knitr::kable(df)
```

Where `timestamp` is the POSIX time format: the number of seconds since midnight
UTC, January 1, 1970. Alternatively a `POSIXct` o `POSIXlt` value can be used.

### Class

The service returns an XML document with several pieces of information:

- The (seed) value represented as a 64 byte (512-bit) hex string value, and its timestamp
- The SHA-512 hash value for the previous record as a 64 byte hex string
- A digital signature (RSA) computed over (in order): version, frequency, timeStamp, seedValue, previousHashValue, errorCode. *Note*: Except for version, the hash is on the byte representations and not the string representations of the data values
- The SHA-512 hash of the signature as a 64 byte hex string
- The service version and its update frequency (currenty, 60 seconds)
- The status of the response
    - 0 = Chain intact, values all good
    - 1 = Start of a new chain of values, previous hash value will be all zeroes
    - 2 = Time between values is greater than the frequency, but the chain is still intact

This package wraps this XML document in the `nrb_record` class, which contains all
the values from the beacon's response.

```{r comment=""}
library(nistrb)
lval <- nrb_get_last()
lval
```


