---
title: "nistr - Quick Reference"
author: "Jesus Castagnetto (<jesus@castagnetto.com>)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nistr - Quick Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package implements R functions and an S3 class to access the
NIST the NIST Randomness Beacon, through its REST web service interface.

The NIST service <q>generates full-entropy bit-strings and posts them in blocks
of 512 bits every 60 seconds</q>^[https://beacon.nist.gov/home]

## Implementation details

### Functions

According to the beacon's documentation, there are five endpoints defined and
this package wraps the call to each one in a function:

- `nrb_get_last()`
    - Endpoint: `https://beacon.nist.gov/rest/record/last`
    - The last generated record
- `nrb_get_next(timestamp)`
    - Endpoint: `https://beacon.nist.gov/rest/record/next/<timestamp>`
    - The record following the given timestamp
- `nrb_get_previous(timestamp)`
    - Endpoint: `https://beacon.nist.gov/rest/record/previous/<timestamp>`
    - The record previous to the given timestamp
- `nrb_get_record(timestamp)`
    - Endpoint: `https://beacon.nist.gov/rest/record/<timestamp>`
    - The record for the given timestamp (or the one closest)
- `nrb_get_start_chain(timestamp)`
    - Endpoint: `https://beacon.nist.gov/rest/record/start-chain/<timestamp>`
    - "The start-of-chain record for the given timestamp

Where `timestamp` is in POSIX time format: the number of seconds since midnight
UTC, January 1, 1970. Alternatively a `POSIXct` o `POSIXlt` value can be used.

### Class

The service returns an XML document with several pieces of information:

- The (seed) value represented as a 64 byte (512-bit) hex string value, and its timestamp
- The SHA-512 hash value for the previous record as a 64 byte hex string
- A digital signature (RSA) computed over (in order): version, frequency, timeStamp, seedValue, previousHashValue, errorCode. *Note*: Except for version, the hash is on the byte representations and not the string representations of the data values
- The SHA-512 hash of the signature as a 64 byte hex string
- The service version and its update frequency (currenty, 60 seconds)
- The status of the response
    - 0 = Chain intact, values all good
    - 1 = Start of a new chain of values, previous hash value will be all zeroes
    - 2 = Time between values is greater than the frequency, but the chain is still intact

This package wraps this XML document in the `nrb_response` class, which contains all
the values from the beacon's response.

An `nrb_response` object contains the properties

- `version`: The current beacon's version
- `frequency`: The beacon's update frequency in seconds
- `url`: The endpoint call
- `timeStamp`: The timestamp (UTC) in Unix Time format
- `seedValue`: The random beacon string for a given timestamp
- `previousOutputValue`: The hash of the previous beacon's record signature
- `signatureValue`: The digital signature of the record
- `outputValue`: The hash of the current signature
- `statusCode`: The service status code for this record
- `statusDesc`: A description of the status code


For example, to get the record generated on Jan 1, 2015 at 10:00h (UTC), we will
use the following code:

```{r}
library(nistrb)
ts <- as.POSIXct("2015-01-01 10:00:00", tz = "UTC")
rec <- nrb_get_record(ts)
rec
```

We can also make sure that the record chain is unbroken, getting the previous
record from the beacon (Jan 1, 2015 09:59h, UTC):

```{r}
prev_ts <- as.POSIXct("2015-01-01 09:59:00", tz = "UTC")
prev_rec <- nrb_get_record(prev_ts)
prev_rec$output == rec$previousOutputValue
```

or the next record:

```{r}
next_ts <- as.POSIXct("2015-01-01 10:01:00", tz = "UTC")
next_rec <- nrb_get_record(next_ts)
rec$output == next_rec$previousOutputValue
```

## Example of use

We can use the NIST Randomness Beacon as a source of reproducible random seeds
that depends on a timestamp, for example:

```{r}
library(nistrb)

# get the number of numeric chars in the seed
# more than a bit contrived as examples go
gen_seed <- function(timestamp) {
  rec <- nrb_get_record(timestamp)
  chrs <- strsplit(rec$seedValue, NULL)
  sum(sapply(chrs, FUN = function(x) {
      grepl(pattern = "[0-9]", x = x)
    }))
}

ts <- as.POSIXct("2015-07-28 00:00:00", tz = "UTC")
gen_seed(ts)

ts <- as.POSIXct("2014-12-25 12:00:00", tz = "UTC")
gen_seed(ts)
```

